---
title: "Outlier_Detection"
author: "Ashwin Agrawal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The function `Outlier_Detection` identifies geographical and statistical outliers based on two different methods known as `AlphaHull` and `Reverse-Jack-Knife`.
The link for the description of above methods are:-
Below is explanation of the function with help of a suitable example using standard GBIF data set.

## Example Data set

The data set used below to explain the function is australian mammals containing 50,000 observations from GBIF portal.

The GBIF R script for downloading above data is shown below.
```{r def1}
library(rgbif)
d1 <- occ_data(
country = "AU",     # Country code for australia
classKey= 359,      # Class code for mammalia
from = 'gbif',
limit=50000,
minimal=FALSE,
hasCoordinate = T
)
#X<-read.csv(file = "C:\\yourfile.csv")
X<-d1$data

knitr::kable(head(X, 10))
```

## Function Code

```{r def2}

Outlier_Detection<-function(X){
  
  ##########################################PART1
  
  DATA<-subset(X,select = c(decimalLongitude,decimalLatitude))
  
  # Alpha Hull approach
  Z<-unique(DATA)
  ahull.obj<-ahull(Z,alpha = 2)
  cat(sprintf("The plot of the alpha Hull is\n"))
  plot(ahull.obj)
  # ashape.obj<-ashape(Z,alpha = 0.2)
  
  # View(DATA)
  # Now checking the points which are in hull using alpha hull and identifying outliers which are outside hull.
  V<-inahull(ahull.obj,p = cbind(DATA$decimalLongitude,DATA$decimalLatitude))
  V<-as.integer(V)
  Val1<-nrow(DATA)
  DATA1<-DATA
  DATA1[,3]<-V
  count<-subset(DATA1,DATA1$V3==0)
  Val2<-nrow(count)
  # This is the condition which checks whether Hull is formed successfully, if not 
  # Reverse Jack knife is used
  if(Val2<=(Val1/2)){
    
    return(DATA1)
    
  }
  ####################################PART2
  else{
    cat(sprintf("The alpha Hull method is not working so using the Reverse-Jack-Knife\n"))
    # Reverse Jack Knife approach
    XXXX1<-rjack((X1$decimalLatitude))
    # XXXX1
    XXXX2<-rjack((X1$decimalLongitude))
    # XXXX2
    
   
    DATA<-as.data.frame(DATA)
    #View(DATA_ALL)
    DATA[,3]<-0
    DATA[XXXX1,3]<-1
    DATA[XXXX2,3]<-1
    
    return(DATA)
  }
  
}

```

## Explanation of the function code
The function code is majorly divided ino two parts:-
PART1:- The Alpha Hull method is used for identifying the outliers by forming a convex hull from coordinates of the points.
If the Alpha Hull method dosen't converge to form a Hull then we go on to the second method.
PART2:- In this part Reverse-Jack-Knife method is used to identify the outliers if Alpha Hull is not working.

Finally the function returns a Data Frame which contains the coordinates (decimalLongitude and decimalLatitude) and a column with flags to represent the outliers.


## Example Walk Through

The australian mammals Data set:-

```{r def3,echo=FALSE}

knitr::kable(head(X, 10))

```

Now using this data to call the `Outlier_Detection` function to get the results.

```{r def4,echo=FALSE,warning=FALSE}
library(alphahull)
library(rgbif)
library(plyr)
library(maptools)
library(raster)
library(biogeo)
X1<-Outlier_Detection(X)
knitr::kable(head(X1, 10))

```

The return value of the function has 3 columns:-
The first and second column contains the cootdinates and the third column contains the flagged value of 0/1, ie if the value is 1 then the observation is outlier and if 0 then it is not an outlier.