---
title: "Taxonomic Flagging and Filtering"
author: "Ashwin Agrawal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

##Taxonomic Data

The taxonomic part of the biodeiversity data deals with the information related to the origin, habitat, chronology and other native information related to specific species. It majorly includes fields like class, kingdom, phylum, genus etc.

## Functions Description and Methodology

The file `Taxonomic_Flagging_and_Filtering` contains functions which help in flagging and filtering data based on taxonnomic fields o fthe data.

The File contains three functions to preform the task.

Function 1:Wrapper for flagging data-`Taxonomic_Flagging`
Simple function to call the main function for flagging data and returning the flagged data frame

Function 2:Wrapper for filtering the flag data-`Taxonomic_Filtering`
Simple function to call the main function for filtering the data returning the filtererd data frame
 
Function 3:Main function:`Taxonomic_Flagging_and_Filtering_main`
This function checks whether the input flag is valid and if so it then flags the data based on the flags and filters the data based on the flag, it returns both flaggedand filtered data frame depending on the user.
Now this function only contains simple flagging anf filtering mechanism based on 
NA values.




## Example Walkthrough

The data set used for explaning the working of the functions is, Australian Mammals is given below:

```{r,def1}

library(rgbif)

d1 <- occ_data(
country = "AU",     # Country code for australia
classKey= 359,      # Class code for mammalia
from = 'gbif',
limit=50000,
minimal=FALSE,
hasCoordinate = T
)

X<-d1$data
#X<-read.csv(file = "C:\\yourfile.csv")
knitr::kable(head(X, 10))

```

Now below is the working of the function based on the above data.

```{r,def2,echo=FALSE}
# Function 1-Flagging Wrapper Function
Taxonomic_Flagging<-function(X,FLAG_NAME){
  
  C<-Taxonomic_Flagging_and_Filtering_main(X,flag_name=FLAG_NAME,flag=TRUE)
  return(C)
  
}

#Function 2-Filtering Wrapper Function

Taxonomic_Filtering<-function(X,FLAG_NAME){
  
  C<-Taxonomic_Flagging_and_Filtering_main(X,flag_name = FLAG_NAME,filter=TRUE)
  return(C)
}


#Function 3-Flagging and Filtering Main function
Taxonomic_Flagging_and_Filtering_main<-function(X,flag_name,filter=FALSE,flag=FALSE){
  
    if(flag_name=="scientificName" || flag_name=="basisOfRecord" || flag_name=="kingdom" || flag_name=="phylum" ||
       flag_name=="order" || flag_name=="genus" || flag_name=="class" || flag_name=="specificEpithet" || 
       flag_name=="institutionCode" || flag_name=="taxonRank" || flag_name=="vernacularName" || flag_name=="name"){
      
      if(flag_name=="scientificName"){
      Data_flag_name<-subset(X,select=c(key,scientificName))
      
      Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$scientificName))
      
      if(flag==TRUE){
#Flag information section
      
      cat(sprintf("The flag convention are as follows\n"))
      cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
      return(Data_flag_name)
      }
      
      
# Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
# and can give user the option to filter the data based on each flag separately  
      
      if(filter==TRUE){
        X$Flag_1<-Data_flag_name$Flag_1
        X<-subset(X,X$Flag_1==0)
        
        drops <- c("Flag_1")
        X<-X[ , !(names(X) %in% drops)]
        
        if(nrow(X)==0){
          stop(paste("The data after filtering is empty"))
        }else{
          cat(sprintf("The filtered data based on scientificName is\n"))
          return(X)
        }
        
          } 
    
      }
      
      
      if(flag_name=="basisOfRecord"){
        Data_flag_name<-subset(X,select=c(key,basisOfRecord))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$basisOfRecord))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on basisOfRecord is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      
      if(flag_name=="kingdom"){
        Data_flag_name<-subset(X,select=c(key,kingdom))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$kingdom))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on scientificName is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      
      if(flag_name=="phylum"){
        Data_flag_name<-subset(X,select=c(key,phylum))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$phylum))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on phylum is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      if(flag_name=="order"){
        Data_flag_name<-subset(X,select=c(key,order))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$order))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on order is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      if(flag_name=="genus"){
        Data_flag_name<-subset(X,select=c(key,genus))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$genus))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on genus is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      if(flag_name=="class"){
        Data_flag_name<-subset(X,select=c(key,class))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$class))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on class is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      
      if(flag_name=="specificEpithet"){
        Data_flag_name<-subset(X,select=c(key,specificEpithet))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$specificEpithet))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on specificEpithet is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      
      if(flag_name=="institutionCode"){
        Data_flag_name<-subset(X,select=c(key,institutionCode))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$institutionCode))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on institutionCode is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      if(flag_name=="taxonRank"){
        Data_flag_name<-subset(X,select=c(key,taxonRank))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$taxonRank))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on taxonRank is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      if(flag_name=="vernacularName"){
        Data_flag_name<-subset(X,select=c(key,vernacularName))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$vernacularName))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on vernacularName is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
      
      if(flag_name=="name"){
        Data_flag_name<-subset(X,select=c(key,name))
        
        Data_flag_name$Flag_1<-as.integer(is.na(Data_flag_name$name))
        
        if(flag==TRUE){
          #Flag information section
          
          cat(sprintf("The flag convention are as follows\n"))
          cat(sprintf("The flag with column name Flag_1 has two values 0/1, if the value is 1 then data field is empty otherwise some data is present\n"))
          return(Data_flag_name)
        }
        
        
        # Now only one flag is developed therefore filter data based on one flag only, in future we can add as many flags as required
        # and can give user the option to filter the data based on each flag separately  
        
        if(filter==TRUE){
          X$Flag_1<-Data_flag_name$Flag_1
          X<-subset(X,X$Flag_1==0)
          
          drops <- c("Flag_1")
          X<-X[ , !(names(X) %in% drops)]
          
          if(nrow(X)==0){
            stop(paste("The data after filtering is empty"))
          }else{
            cat(sprintf("The filtered data based on name is\n"))
            return(X)
          }
          
        } 
        
      }
      
      
    }else{
          stop(paste("The flag is not present"))
      
    }  
                           
}


```


The results for only the flagging part are:

```{r,def3}
Z1<-Taxonomic_Flagging(X,FLAG_NAME = "name") 
knitr::kable(head(Z1, 10))

```

The results after and flagging and filtering the data based on the flags are:

```{r,def4}
Z2<-Taxonomic_Filtering(X,FLAG_NAME = "name")
knitr::kable(head(Z2, 10))
```


